Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-30 ind_files/05_unification_discussion.tex
Make notation more compact

Equation \eqref{nD_eqn_for_frequencies}, which we derived in chapter \ref{chap_BD}, is a very general equation for how frequencies change over time in stochastic populations. To recap, we started with a population which can contain up to $m$ different (fixed) types of individuals. The population as a whole is characterized by a vector $\mathbf{n}(t) = [n_1(t), \ldots, n_m(t)]$ indexing the number of individuals of each type in the population. Changes of the population are through either birth or death of single individuals. On ecological grounds, we postulated the existence of a `system-size' parameter $K>0$ that leads to density-dependent growth and prevents the population from growing infinitely large. We then moved from numbers $\mathbf{n}(t)$ to population densities $\mathbf{x}(t)=\mathbf{n}/K$ by assuming that birth and death rates depend on population densities $\mathbf{x}$ and not only on population numbers $\mathbf{n}$. In the regime where $K$ is not too small (corresponding to `medium sized' populations), we found a continuous description of how $\mathbf{x}$ changes stochastically. We further assumed that each type has a per-capita birth rate $b^{\textrm{(ind)}}(\mathbf{x})$, a per-capita death rate $d^{\textrm{(ind)}}(\mathbf{x})$, and an additional term $\mu Q_{i}(\mathbf{x})$ representing additional birth of type $i$ individuals due to mutations of other types. All three of these functions may in general vary in an arbitrarily complicated frequency-dependent manner (or be frequency independent constants, of course). We then moved from type densities $\mathbf{x}(t)$ to type frequencies $\mathbf{p}(t)$ and saw that $p_i$, the frequency of the $i\textsuperscript{th}$ type in the population, obeys the stochastic differential equation:

@@

where $N_K = \sum x_i$ is the total population size scaled by $K$ (and thus $KN_K$ is the total population size), $w_i(\mathbf{x}) = b^{\textrm{(ind)}}_{i}(\mathbf{x}) - d^{\textrm{(ind)}}_{i}(\mathbf{x})$ and $\tau_i(\mathbf{x}) = b^{\textrm{(ind)}}_{i}(\mathbf{x}) + d^{\textrm{(ind)}}_{i}(\mathbf{x})$ are respectively the Malthusian fitness and per-capita turnover rate of the $i\textsuperscript{th}$ type, and $A_i^{+} \coloneqq x_i\tau_i(\mathbf{x}) + \mu Q_i(\mathbf{x})$. Each $W^{(i)}_t$ is an independent one-dimensional Wiener process (standard Brownian motion). Letting $K \to \infty$  in \eqref{nD_stochastic_RM} recovers the replicator-mutator equation in the infinite population limit. The first term of \eqref{nD_stochastic_RM} represents the direct effects of forces captured in classic deterministic models, and reflects a selection-mutation balance. However, finite populations experience a new directional force dependent on $\tau_i(\mathbf{x})$, the per-capita turnover rate of type $i$, that cannot be captured in infinite population models~\citep{kuosmanen_turnover_2022}. Remarkably, this term acts in a way that is mathematically identical to the classical action of selection and mutation in infinite population models as captured by the first term in \eqref{nD_stochastic_RM}, but in the opposite direction - A higher relative $\tau_i$ leads to a decrease in frequency (Notice the minus sign before the second term in \eqref{nD_stochastic_RM}).

@@

We can also calculate how the statistical mean value of any type-level quantity (see section \ref{sec_stat_measures}) changes over time. Let $f$ be any type level quantity, with value $f_i(t)$ for the $i\textsuperscript{th}$ type. I allow for the possibility of $f_i$ to vary over time. By multiplying both sides of
equation \eqref{nD_stochastic_RM} by $f_i$ and summing over all $i$ (The same steps as going from \eqref{nD_replicator_mutator} to \eqref{nD_Price_time_dependent}), we see that the statistical mean $\overline{f}$ of the quantity in the population varies as:
\begin{equation}
\label{nD_stochastic_Price}
\begin{aligned}
%The \vphantom is to vertically align the texts under the \underbraces
d\overline{f} &= {\underbrace{\vphantom{ \frac{1}{KN_K(t)} }\textrm{Cov}(w,f)dt}_{\substack{\text{Classical} \\ \text{selection}}}} \ - \ {\underbrace{\frac{1}{KN_K(t)}\textrm{Cov}(\tau,f)dt}_{\substack{\text{Noise-induced} \\ \text{selection}}}} \ +  \  {\underbrace{\vphantom{ \frac{1}{KN_K(t)} }\overline{\left(\frac{\partial f}{\partial t}\right)}dt}_{\substack{\text{Ecological timescale} \\ \text{feedbacks due to} \\ \text{time-dependence of $f_i$}}}}\\[15pt]
&+ \underbrace{\mu\left(1-\frac{1}{KN_K(t)}\right)\left(\sum\limits_{i=1}^{m}f_iQ_i(\mathbf{p}) - \overline{f}\sum\limits_{i=1}^{m}Q_i(\mathbf{p})\right)dt}_{\text{Transmission bias/mutational effects}}\\[15pt]
&+ \underbrace{\frac{1}{\sqrt{K}N_{K}(t)}\left(\sum\limits_{i=1}^{m}\left(f_i-\overline{f}\right)\sqrt{A_i^+}dW_{t}^{(i)}\right)}_{\text{Stochastic fluctuations}}
\end{aligned} 
\end{equation}
where all covariances are understood in the statistical sense (Note that since $w_i$, $\tau_i$, $\overline{w}$, and $\overline{\tau}$ are stochastic processes depending on $\mathbf{p}$, the terms $\textrm{Cov}(w,f)$ and $\textrm{Cov}(\tau,f)$ are themselves stochastic processes; See section \ref{sec_stat_measures}). Taking $K \to \infty$ in equation \eqref{nD_stochastic_Price} recovers the standard Price equation as the infinite population limit (either \eqref{nD_Price_time_dependent} or \eqref{nD_Price} based on whether $f_i$ varies with time; also see~\cite{rice_universal_2020} for a stochastic Price equation in a discrete-time setting).

@@

The third term of \eqref{nD_stochastic_Price} is relevant in both finite and infinite populations whenever $f_i$ can vary over time and represents feedback effects on the quantity $f_i$ of the $i\textsuperscript{th}$ species over short (`ecological') time-scales. Such feedback is often through a changing environment or phenotypic/behavioral plasticity, but other biological phenomena may also be at play. The fourth term of \eqref{nD_stochastic_Price} is a transmission bias term, with a correction factor due to noise-induced selection. Finally, the last term of \eqref{nD_stochastic_Price} describes the role of stochastic fluctuations. The contributions of this last term are `directionless' due to the $dW_t$ factors, and this term vanishes when we take a conditional expectation value over the underlying probability space. I will denote this probabilistic expectation value operation by $\mathbb{E}[\cdot]$ to distinguish it from the statistical mean $\overline{f}$ of a type-level quantity in the population (see section \ref{sec_stat_measures} for a brief discussion on the difference between statistical averages and probabilistic expectations in this context). Note that this expectation is conditioned on the initial state of the population, and thus $\mathbb{E}[\cdot]$ is really shorthand for $\mathbb{E}[\ \cdot \ | \ \mathbf{X}_0 = \mathbf{x}_0]$.

Two particularly interesting implications of \eqref{nD_stochastic_Price} are realized upon ignoring mutations by setting $\mu = 0$ and then substituting $f=w$ and $f = \tau$. We first note that:
\begin{align}
\textrm{Cov}(w,\tau) &=\textrm{Cov}\left( b^{\textrm{(ind)}}(\mathbf{x}) - d^{\textrm{(ind)}}(\mathbf{x}) \ , \   b^{\textrm{(ind)}}(\mathbf{x}) + d^{\textrm{(ind)}}(\mathbf{x})\right)\\
&= \sigma^2_{b^{\textrm{(ind)}}(\mathbf{x})} - \sigma^2_{d^{\textrm{(ind)}}(\mathbf{x})}\label{nD_cross_covariance}
\end{align}
It is important to remember once again that just like the statistical mean, the statistical variance $\sigma^2_{f}(t)$ of a type-level quantity $f$ is a random variable obtained by calculating the variance of the quantity \emph{in the population} at time $t$, and is not to be confused with the probabilistic/ensemble variance obtained by calculating the variance of a quantity \emph{over different realizations} of the stochastic process (see section \ref{sec_stat_measures}). Upon substituting $f = w$ in \eqref{nD_stochastic_Price} and taking expectations over the underlying probability space, we obtain:
\begin{align}
\label{nD_stochastic_Fisher}
%The \vphantom is to vertically align the texts under the underbraces of different terms
\mathbb{E}\left[\frac{d\overline{w}}{dt}\right] &= 
\negmedspace {\underbrace{\mystrut{4ex} \ \vphantom{ \frac{\sigma^2_{b^{\textrm{(ind)}}} - \sigma^2_{d^{\textrm{(ind)}}}}{KN_K(t)} } \mathbb{E}\left[\sigma^2_{w}\right] \ }_{\substack{\text{Fisher's} \\ \text{Fundamental theorem}}}} \ - \ {\underbrace{\mystrut{4ex}\mathbb{E}\left[\frac{\sigma^2_{b^{\textrm{(ind)}}} - \sigma^2_{d^{\textrm{(ind)}}}}{KN_K(t)}\right]}_{\substack{\text{Noise-induced} \\ \text{selection}}}} \ + {\underbrace{\mystrut{4ex} \vphantom{ \frac{\sigma^2_{b^{\textrm{(ind)}}} - \sigma^2_{d^{\textrm{(ind)}}}}{KN_K(t)} } \mathbb{E}\left[\hphantom{a}\overline{\hphantom{a}\frac{\partial w}{\partial t}\hphantom{a}}\hphantom{a}\right]}_{\substack{\text{Short-term (ecological)} \\ \text{feedbacks to fitness}}}}
\end{align}
Taking $K \to \infty$ in \eqref{nD_stochastic_Fisher} recovers a well-known equation in population genetics upon noting that the process tends to a deterministic process as $K \to \infty$, as noted in section \ref{sec_nD_det_limit}, and thus the expectation value in the infinite population case is superfluous. The first term, $\sigma^2_w$, is the subject of Fisher's fundamental theorem~\citep{fisher_genetical_1930,  price_fishers_1972, frank_fishers_1992, kokko_stagnation_2021}. The second term of equation \eqref{nD_stochastic_Fisher} is a manifestation of noise-induced selection and vanishes in the infinite population limit, and is thus particular to finite populations. Finally, the last term arises in both finite and infinite populations whenever $w_i$ can vary over time~\citep{frank_fishers_1992,kokko_stagnation_2021,baez_fundamental_2021}, be it through frequency-dependent selection, phenotypic plasticity, varying environments, or other ecological mechanisms, and represents feedback effects on the fitness $w_i$ of the $i\textsuperscript{th}$ species over short (`ecological') time-scales. The fact that~\citet{fisher_genetical_1930} appears to have been rather vague and dismissive of this feedback has led to much discussion, debate, and confusion about the interpretation, importance, and implications of his `fundamental theorem' (see~\cite{kokko_stagnation_2021} and sources cited therein).

Carrying out the same steps with $f = \tau$ in \eqref{nD_stochastic_Price} yields a new equation due to~\citet{kuosmanen_turnover_2022}. The result is an analog of Fisher's fundamental theorem for the turnover rates, and reads:
\begin{equation}
\label{nD_stochastic_Fisher_turnover}
\mathbb{E}\left[\frac{d\overline{\tau}}{dt}\right] = \underbrace{\vphantom{ \mathbb{E}\left[\frac{\sigma^2_{\tau}}{KN_K(t)}\right] } \mathbb{E}\left[\sigma^2_{b^{\textrm{(ind)}}} - \sigma^2_{d^{\textrm{(ind)}}}\right]}_{\substack{\text{Classical selection} \\ \text{effects}}} \ - \ \underbrace{ \mathbb{E}\left[\frac{\sigma^2_{\tau}}{KN_K(t)}\right]}_{\substack{\text{Noise-induced selection} \\ \text{effects}}} \ + \ \underbrace{ \vphantom{ \mathbb{E}\left[\frac{\sigma^2_{\tau}}{KN_K(t)}\right] } \mathbb{E}\left[\hphantom{a}\overline{\hphantom{a}\frac{\partial \tau}{\partial t}\hphantom{a}}\hphantom{a}\right]}_{\substack{\text{Short-term (ecological)} \\ \text{feedbacks to $\tau_i$}}}
\end{equation}
The implications of this equation have been extensively discussed in~\cite{kuosmanen_turnover_2022}, which is where I refer the interested reader.

>>>>

Equation \eqref{nD_eqn_for_frequencies}, which we derived in chapter \ref{chap_BD}, is a very general equation for how frequencies change over time in stochastic populations. To recap, we started with a population which can contain up to $m$ different (fixed) types of individuals. The population as a whole is characterized by a vector $\mathbf{n}(t) = [n_1(t), \ldots, n_m(t)]^{\mathrm{T}}$ indexing the number of individuals of each type in the population. Changes of the population are through either birth or death of single individuals. On ecological grounds, we postulated the existence of a `system-size' parameter $K>0$ that leads to density-dependent growth and prevents the population from growing infinitely large. We then moved from numbers $\mathbf{n}(t)$ to population densities $\mathbf{x}(t)=\mathbf{n}/K$ by assuming that birth and death rates depend on population densities $\mathbf{x}$ and not only on population numbers $\mathbf{n}$. In the regime where $K$ is not too small (corresponding to `medium sized' populations), we found a continuous description of how $\mathbf{x}$ changes stochastically. We further assumed that each type has a per-capita birth rate $b^{\textrm{(ind)}}(\mathbf{x})$, a per-capita death rate $d^{\textrm{(ind)}}(\mathbf{x})$, and an additional term $\mu Q_{i}(\mathbf{x})$ representing additional birth of type $i$ individuals due to mutations of other types. All three of these functions may in general vary in an arbitrarily complicated frequency-dependent manner (or be frequency independent constants, of course). We then moved from type densities $\mathbf{x}(t)$ to type frequencies $\mathbf{p}(t)$ and saw that $p_i$, the frequency of the $i\textsuperscript{th}$ type in the population, obeys the stochastic differential equation:

@@

where $N_K = \sum x_i = \sum n_i /K$ is the total population size scaled by $K$ (and thus $KN_K$ is the total population size), $w_i(\mathbf{x}) = b^{\textrm{(ind)}}_{i}(\mathbf{x}) - d^{\textrm{(ind)}}_{i}(\mathbf{x})$ and $\tau_i(\mathbf{x}) = b^{\textrm{(ind)}}_{i}(\mathbf{x}) + d^{\textrm{(ind)}}_{i}(\mathbf{x})$ are respectively the Malthusian fitness and per-capita turnover rate of the $i\textsuperscript{th}$ type, and $A_i^{+} \coloneqq x_i\tau_i(\mathbf{x}) + \mu Q_i(\mathbf{x})$. Each $W^{(i)}_t$ is an independent one-dimensional Wiener process (standard Brownian motion). Letting $K \to \infty$  in \eqref{nD_stochastic_RM} recovers the replicator-mutator equation in the infinite population limit. The first term of \eqref{nD_stochastic_RM} represents the direct effects of forces captured in classic deterministic models, and reflects a selection-mutation balance. However, finite populations experience a new directional force dependent on $\tau_i(\mathbf{x})$, the per-capita turnover rate of type $i$, that cannot be captured in infinite population models~\citep{kuosmanen_turnover_2022}. Remarkably, this term acts in a way that is mathematically identical to the classical action of selection and mutation in infinite population models as captured by the first term in \eqref{nD_stochastic_RM}, but in the opposite direction - A higher relative $\tau_i$ leads to a decrease in frequency (Notice the minus sign before the second term in \eqref{nD_stochastic_RM}).

@@

We can also calculate how the statistical mean value of any type-level quantity (see section \ref{sec_stat_measures}) changes over time. Let $f$ be any type level quantity, with value $f_i(t)$ for the $i\textsuperscript{th}$ type. I allow the value $f_i$ to possibly vary over time (for example, due to plasticity or a changing environment). By multiplying both sides of
equation \eqref{nD_stochastic_RM} by $f_i$ and summing over all $i$ (The same steps as going from \eqref{nD_replicator_mutator} to \eqref{nD_Price_time_dependent}), we see that the statistical mean $\overline{f}$ of the quantity in the population varies as:
\begin{equation}
\label{nD_stochastic_Price}
%The \vphantom is to vertically align the underbraces
\resizebox{\textwidth}{!}{$\displaystyle d\overline{f} = {\underbrace{\vphantom{ \frac{1}{KN_K(t)} }\textrm{Cov}(w,f)dt}_{\substack{\text{Classical} \\ \text{selection}}}} - {\underbrace{\frac{1}{KN_K(t)}\textrm{Cov}(\tau,f)dt}_{\substack{\text{Noise-induced} \\ \text{selection}}}} +  {\underbrace{\vphantom{ \frac{1}{KN_K(t)} }\overline{\left(\frac{\partial f}{\partial t}\right)}dt}_{\substack{\text{Ecological} \\ \text{feedbacks}}}}
+ \underbrace{\vphantom{ \frac{1}{KN_K(t)} }M_{\overline{f}}(\mathbf{p},N_K)dt}_{\substack{\text{Mutational} \\ \text{effects}}}
+ \underbrace{\frac{1}{\sqrt{K}N_{K}(t)}dW_{\overline{f}}}_{\substack{\text{Stochastic} \\ \text{fluctuations}}}$}
\end{equation}
where all covariances are understood in the statistical sense (see section \ref{sec_stat_measures}). Here,
\begin{equation}
	M_{\overline{f}}(\mathbf{p},N_K) \coloneqq \mu\left(1-\frac{1}{KN_K(t)}\right)\left(\sum\limits_{i=1}^{m}f_iQ_i(\mathbf{p}) - \overline{f}\sum\limits_{i=1}^{m}Q_i(\mathbf{p})\right)
\end{equation}
is a term capturing mutational effects/transmission fidelity that vanishes in the low mutation rate ($\mu \to 0$) limit, and 
\begin{equation}
	dW_{\overline{f}} \coloneqq \sum\limits_{i=1}^{m}\left(f_i-\overline{f}\right)\sqrt{A_i^+}dW_{t}^{(i)}
\end{equation}
is a stochastic integral term describing un-directed stochastic fluctuations and vanishes upon taking expectation values. Note that the $1/\sqrt{K}$ factor outside this term in equation \eqref{nD_stochastic_Price} means that these stochastic fluctuations also disappear in the infinite population ($K \to \infty$) limit.

Taking $K \to \infty$ in equation \eqref{nD_stochastic_Price} recovers the standard Price equation as the infinite population limit (either \eqref{nD_Price_time_dependent} or \eqref{nD_Price} based on whether $f_i$ varies with time; also see~\cite{rice_universal_2020} for a stochastic Price equation in a discrete-time setting).

@@

The third term of \eqref{nD_stochastic_Price} is relevant in both finite and infinite populations whenever $f_i$ can vary over time and represents feedback effects on the quantity $f_i$ of the $i\textsuperscript{th}$ species over short (`ecological') time-scales. Such feedback could be through a changing environment, phenotypic/behavioral plasticity, or any manner of other `ecological' phenomena. This is the term that captures eco-evolutionary feedback loops.

The fourth term of \eqref{nD_stochastic_Price} is a transmission bias term, with a correction factor due to noise-induced selection. Finally, the last term of \eqref{nD_stochastic_Price} describes the role of stochastic fluctuations. The contributions of this last term are `directionless' due to the $dW_t$ factors, and this term vanishes when we take a conditional expectation value over the underlying probability space. We denote this probabilistic expectation value operation by $\mathbb{E}[\cdot]$ to distinguish it from the statistical mean \eqref{nD_mean}. Note that this expectation is conditioned on the initial state of the population, and thus $\mathbb{E}[\cdot]$ is really shorthand for $\mathbb{E}[\ \cdot \ | \ \mathbf{X}_0 = \mathbf{x}_0]$.

Two particularly interesting implications of \eqref{nD_stochastic_Price} are realized upon ignoring mutations by setting $\mu = 0$ and then substituting $f=w$ and $f = \tau$. We first note that:
\begin{align}
\textrm{Cov}(w,\tau) &=\textrm{Cov}\left( b^{\textrm{(ind)}}(\mathbf{x}) - d^{\textrm{(ind)}}(\mathbf{x}) \ , \   b^{\textrm{(ind)}}(\mathbf{x}) + d^{\textrm{(ind)}}(\mathbf{x})\right)\\
&= \sigma^2_{b^{\textrm{(ind)}}(\mathbf{x})} - \sigma^2_{d^{\textrm{(ind)}}(\mathbf{x})}\label{nD_cross_covariance}
\end{align}
It is important to remember once again that just like the statistical mean, the statistical variance $\sigma^2_{f}(t)$ of a type-level quantity $f$ is a random variable obtained by calculating the variance of the quantity \emph{in the population} at time $t$, and is not to be confused with the probabilistic/ensemble variance obtained by calculating the variance of a quantity \emph{over different realizations} of the stochastic process (see section \ref{sec_stat_measures}). Upon substituting $f = w$ in \eqref{nD_stochastic_Price} and taking expectations over the underlying probability space, we obtain:
\begin{align}
\label{nD_stochastic_Fisher}
%The \vphantom is to vertically align the texts under the underbraces of different terms
\mathbb{E}\left[\frac{d\overline{w}}{dt}\right] &= 
\negmedspace {\underbrace{\mystrut{4ex} \ \vphantom{ \frac{\sigma^2_{b^{\textrm{(ind)}}} - \sigma^2_{d^{\textrm{(ind)}}}}{KN_K(t)} } \mathbb{E}\left[\sigma^2_{w}\right] \ }_{\substack{\text{Fisher's} \\ \text{fundamental} \\ \text{theorem}}}} \ - \ {\underbrace{\mystrut{4ex}\mathbb{E}\left[\frac{\sigma^2_{b^{\textrm{(ind)}}} - \sigma^2_{d^{\textrm{(ind)}}}}{KN_K(t)}\right]}_{\substack{\text{Noise-induced} \\ \text{selection}}}} \ + {\underbrace{\mystrut{4ex} \vphantom{ \frac{\sigma^2_{b^{\textrm{(ind)}}} - \sigma^2_{d^{\textrm{(ind)}}}}{KN_K(t)} } \mathbb{E}\left[\hphantom{a}\overline{\hphantom{a}\frac{\partial w}{\partial t}\hphantom{a}}\hphantom{a}\right]}_{\substack{\text{Eco-evolutionary} \\ \text{feedbacks to fitness}}}}
\end{align}
Taking $K \to \infty$ in \eqref{nD_stochastic_Fisher} recovers a well-known equation in population genetics upon noting that the process tends to a deterministic process as $K \to \infty$, as noted in section \ref{sec_nD_det_limit}, and thus the expectation value in the infinite population case is superfluous. The first term, $\sigma^2_w$, is the subject of Fisher's fundamental theorem~\citep{fisher_genetical_1930,  price_fishers_1972, frank_fishers_1992, kokko_stagnation_2021}. The second term of equation \eqref{nD_stochastic_Fisher} is a manifestation of noise-induced selection and vanishes in the infinite population limit, and is thus particular to finite populations. Finally, the last term arises in both finite and infinite populations whenever $w_i$ can vary over time~\citep{frank_fishers_1992,kokko_stagnation_2021,baez_fundamental_2021}, be it through frequency-dependent selection, phenotypic plasticity, varying environments, or other ecological mechanisms, and represents feedback effects on the fitness $w_i$ of the $i\textsuperscript{th}$ species over short (`ecological') time-scales. The fact that~\citet{fisher_genetical_1930} appears to have been rather vague and dismissive of this feedback has led to much discussion, debate, and confusion about the interpretation, importance, and implications of his `fundamental theorem' (see~\cite{kokko_stagnation_2021} and sources cited therein).

Carrying out the same steps with $f = \tau$ in \eqref{nD_stochastic_Price} yields a new equation due to~\citet{kuosmanen_turnover_2022}. The result is an analog of Fisher's fundamental theorem for the turnover rates, and reads:
\begin{equation}
\label{nD_stochastic_Fisher_turnover}
\mathbb{E}\left[\frac{d\overline{\tau}}{dt}\right] = \underbrace{\vphantom{ \mathbb{E}\left[\frac{\sigma^2_{\tau}}{KN_K(t)}\right] } \mathbb{E}\left[\sigma^2_{b^{\textrm{(ind)}}} - \sigma^2_{d^{\textrm{(ind)}}}\right]}_{\substack{\text{Classical selection} \\ \text{effects}}} \ - \ \underbrace{ \mathbb{E}\left[\frac{\sigma^2_{\tau}}{KN_K(t)}\right]}_{\substack{\text{Noise-induced selection} \\ \text{effects}}} \ + \ \underbrace{ \vphantom{ \mathbb{E}\left[\frac{\sigma^2_{\tau}}{KN_K(t)}\right] } \mathbb{E}\left[\hphantom{a}\overline{\hphantom{a}\frac{\partial \tau}{\partial t}\hphantom{a}}\hphantom{a}\right]}_{\substack{\text{Eco-evolutionary} \\ \text{feedbacks to $\tau_i$}}}
\end{equation}
The implications of this equation have been extensively discussed in~\cite{kuosmanen_turnover_2022}, which is where I refer the interested reader.
